Master thesis problem statement: Keyword search over heterogeneous sources

### 2012-10-11 doc/thesis/latex/das_logs.tex
updated problem definition, solutions and literature review. review to be finished

\documentclass[a4paper,11pt,draft]{article}
\special{papersize=210mm,297mm}

% twocolumn
%\usepackage{fullpage} 
\usepackage[cm]{fullpage}
%\usepackage[margin=0.5in]{geometry}
%\usepackage{times}
\usepackage{color}
%\usepackage[small,compact]{titlesec}

\bibliographystyle{alpha}

% usually first paragraph is NOT indented, so this is commented out: 
% \usepackage{indentfirst}

\begin{document}

% Article top matter
\title{Master thesis problem statement: Keyword search over heterogeneous sources} %\LaTeX is a macro for printing the Latex logo
\author{Vidmantas Zemleris}  %\texttt formats the text to a typewriter style font
\date{\today}  %\today is replaced with the current date

%\maketitle

%
 \centerline{\Large \bf Searching heterogeneous data-sources: Master thesis problem statement} %% Paper title
 \medskip

 \centerline{Vidmantas Zemleris, \today}  %% Author name(s)
 \medskip

\section{Introduction}
At large scientific collaborations like the CMS Experiment at CERN's LHC that includes more than 3000 collaborators data usually resides on a fair number of autonomous and heterogeneous proprietary systems each serving it's own purpose
\footnote{For instance, at CERN,  due to many reasons (e.g. research and need of freedom, politics of institutes involved) software projects usually evolve in independent fashion resulting in fair number of proprietary systems\cite{Koch00CERN}. Further high turnover makes it harder extending these systems}. As data stored on one system may be related to data residing on other systems%
	\footnote{For example, datasets containing physics events are registered at DBS, while the physical location of files is tracked by Phedex which also takes care of their transfers within the worldwide grid storage}%
, users are in need of a centralized and easy-to-use solution for locating and combining data from all these multiple services.

Using a highly structured language like SQL is problematic because users need to know not only the language but also where to find the information and also lots of technical details like schema. A data integration system based on simple structured queries is already in place. Still {\color{red}various improvements including support} for less restricted keyword queries and improvements to system's usability and performance still have to be researched.

% {\color{red}A remarkable opportunity we have is possibility to test ideas on a sample of ~3000 people working at CMS.}

\section{Case study: the CMS Experiment at CERN}
Users' information need may vary greatly depending on their role, however most of the time they are interested in locating a \textit{full set} of entities matching some selection criteria, e.g.:
%\footnote{in contrary to exploratory approach where  a subset of best matches is enough}
   \begin{itemize}
  		\item find all \textit{files} from \textit{dataset(s)} matching wild-card query each containing some of the 'interesting' \textit{runs} from a list provided (Release validation teams)
         \item find (all) \textit{datasets} related some specific physics phenomena\footnote{In case of dataset this data is present in filename or run} together with conditions describing how this data was recorded by detector or simulated which are present in separate autonomous system than the datasets (Physicists)
         % TODO:is this a good example? conditions is a separate query now...
         \item find all \textit{datasets} matching some pattern stored at a given \textit{site} (filtering on entities stored on separate services)
   \end{itemize}                		
% In such dynamic environment it is preferred to choose local-as-view approach for information integration instead of creating a global schemata\cite{Koch00CERN}.

For more use-cases of data retrieval at CMS Experiment see \cite{CMS_data08}.

%Following a similar approach a 
\subsection*{The Data Aggregation System}

The Data Aggregation System (DAS)\cite{Kuznetsov2010, Kuznetsov2011} was created which allows integrated access to a number of proprietary data-sources by processing user's queries on demand - it queries the data-sources, merges the results, and caches them for subsequent use. DAS uses \textit{Boolean retrieval model} as users are often interested in retrieving ALL the items matching their query.

Currently the queries specify what entity the user is interested in (dataset, file, etc) and provide selection criteria (attribute=value, name BETWEEN [v1, v2]) operators. The combined query results could be later 'piped' for further filtering and aggregation (min, avg, etc), e.g.:

{\footnotesize 
\begin{verbatim}
dataset=*RelVal* | grep dataset.nevents >1000 | avg(dataset.size), median(dataset.size)
\end{verbatim}
}

The query above would return average and median datasets sizes  of ones containing  \textit{RelVal} in their name having more than 1000 events.

Queries could be run either from web browser or through  command line interface where the results could  be fed into another application (e.g. program doing physics analysis or automatic software release validation).

% \textbf{\color{red}Examples of more general uses: Digital libraries, Stock markets, [ Weather, Flight costs]}

\section{Problem statement}

DAS is based on \textit{Virtual Integration} where data is left at the sources allowing data to be volatile (e.g. new records gathered or existing records updated; service descriptions may be changed; new services added), while some subsets of the data could be fairly static (e.g. datasets stored at DBS do not change often). 
Further as the total data that is being queried could be fairly large (with growth of order of 1TB per year) and the data-services may have certain constraints (e.g. only certain queries allowed) as their main goal is supporting production (data taking from the CMS detector).

Therefore, it is important to balance between returning locally cached query results quickly   and between issuing queries to slow data-services to get up to date results.
It is also not less important to provide users with easy to use interface having a fairly flat learning curve, while at the same time not loosing the functionality of fairly complex queries. % expressive power.
\textbf{\color{red}Another issue is about handling distributed queries, where either selection criteria or selection values could come from distributed sources -- the biggest problem could be in the APIs they [DO NOT] provide}.

%TODO: Further it has to comply with dataset constraints

% \textbf{\color{red}DO WE also have proprietary big databases that could be useful to be searched, but the standard integration work would be to heavy (or sub-parts of existing systems that are not covered by APIs, and could be accessible through DBs directly -- these potentially more expensive queries could processed with lower priority)}

\subsection{Ease of use: Query Language over the mediated schema}
	Even a very simple structured query language that also contain entity names over the mediated schema may seem {\color{red}hard} to learn, especially in the beginning.
   On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} (even though there may exist slight differences in their naming on different data-sources). 
   So some sort of guiding helping user to build the query shall be useful.

   Supporting non-structured keyword queries is also worth investigation as many users reporting they are missing this Google-like search experience. {\color{red}Because DAS uses Boolean IR we could only rank specific structured queries.}

	Further a minimum set of search predicates is imposed by APIs (mainly because of performance reasons) and user has to be at least informed what is he expected to provide.

Also to consider: input ambiguity and typos, some queries are more common than others.

\subsection{Performance}

\begin{itemize}
	     					\item  Query prioritization: {\color{red}maybe we could have prioritization in general, the heavier your query is the more you have to wait in favour of  the light queries}: 
	     					(we could have some sort of query cost evaluation based on history or manually predefined scores per API)

                		\item Explore more intelligent caching

	                		% TODO: \item query rewriting then used with grep=smf then the same item is available as selection key. check how many instances.

					\item Could displaying Partial Results improve performance?
					\begin{itemize}
					     \item given current APIs to do aggregation over data sources, DAS has to fetch ALL records matching the query instead of only the first page.
%                        \item ??? do we have cases then multiple APIs of the same service could be called for the same data, can we choose just one???
                        \item {\color{red}e.g. intermediary aggregation results while still calculating}

					\end{itemize}

	                	 \item scale testing - if we are storing lots of historical info. MongoDB is not so performant if DB cant fit in memory. One of well known solutions is installing SSD.                        
\end{itemize}

\subsection{Handling distributed search efficiently}

\pagebreak
\include{proposed_solutions}

\pagebreak
\include{literature_review}

\section{Work status}
TODO

\begin{verbatim}
- obtained DB copy of biggest data provider DBS (currently 80 GB + 200 GB indexes)
- preliminary literature review (to be continued deeper + waiting for book arrival: Principles of Data Integration)

--- some analysis of DAS logs

Upcomming Work items:
- couple of fairly simple prototypes of UI/access patters for simpler DAS querying
- check on performance improvements after creation of materialized view(s)
- interview (more) DAS users
\end{verbatim}

% ---  Bibliography
\bibliographystyle{unsrt}
\addcontentsline{toc}{chapter}{Bibliography}

%\nocite{*}
\thispagestyle{empty}
\bibliography{refs}

\pagebreak
\appendix
\section{DAS Query logs}

non parsible queries: 98923

\subsection*{Most common query patterns (Feb 2012-June 2012)}
\begin{verbatim}
total valid queries: 569408

50.16% (285605) :	dataset dataset.name=?
13.54% (77071)  :	site dataset.name=?
8.96%  (51035) 	:	file dataset.name=?
5.88%  (33504) 	:	run dataset.name=?
2.59%  (14739) 	:	release dataset.name=?
2.11%  (12036) 	:	config dataset.name=?
1.65%  (9420) 	:	dataset run.run_number=?
1.34%  (7642) 	:	block dataset.name=?
1.24%  (7084) 	:	dataset site.name=?
1.15%  (6562) 	:	dataset dataset.name=? release.name=?
1.10%  (6287) 	:	parent dataset.name=?
0.96%  (5488) 	:	file file.name=?
0.92%  (5245) 	:	dataset dataset.name=? status.name=?
0.77%  (4363) 	:	dataset release.name=?
0.75%  (4257) 	:	file dataset.name=? run.run_number=?
0.68%  (3874) 	:	run run.run_number=?
0.53%  (2994) 	:	site site.name=?
0.45%  (2576) 	:	site file.name=?
0.45%  (2556) 	:	dataset file.name=?
0.43%  (2438) 	:	lumi file.name=?
0.40%  (2282) 	:	dataset dataset.name=? site.name=?
0.35%  (1999) 	:	file block.name=?
0.35%  (1970) 	:	dataset dataset.name=? run.run_number=?
0.29%  (1640) 	:	group dataset.name=?
0.29%  (1631) 	:	lumi run.run_number=?
0.25%  (1402) 	:	child dataset.name=?
0.22%  (1268) 	:	run file.name=?
0.21%  (1204) 	:	block block.name=?
0.19%  (1088) 	:	release release.name=?
0.16%  (936) 	:	site block.name=?
0.12%  (703) 	:	file dataset.name=? lumi.number=? run.run_number=?
0.10%  (594) 	:	parent file.name=?
0.08%  (455) 	:	dataset block.name=?
0.06%  (365) 	:	dataset dataset.name=? datatype.name=? release.name=?
0.06%  (360) 	:	file dataset.name=? site.name=?

Interesting non structured queries:
T!_CERN
*herwig*/AODSIM

\end{verbatim}
% py dataset=/DYJetsToLL_M-50_TuneZ2Star_8TeV-madgraph-tarball/Summer12_DR53X-PU_S10_START53_V7A-v1/AODSIM

\end{document}

>>>>

\appendix
\section{DAS Query logs}

\subsection*{Most common query patterns (from 2011-06-21 to 2012-10-01)}
\begin{verbatim}
total valid queries: 569,408
not well formed queries (e.g. free text or with typing mistakes): 98,923

50.16% (285605):	dataset dataset.name=?
13.54% (77071) :	site dataset.name=?
8.96%  (51035) :	file dataset.name=?
5.88%  (33504) :	run dataset.name=?
2.59%  (14739) :	release dataset.name=?
2.11%  (12036) :	config dataset.name=?
1.65%  (9420) 	:	dataset run.run_number=?
1.34%  (7642) 	:	block dataset.name=?
1.24%  (7084) 	:	dataset site.name=?
1.15%  (6562) 	:	dataset dataset.name=? release.name=?
1.10%  (6287) 	:	parent dataset.name=?
0.96%  (5488) 	:	file file.name=?
0.92%  (5245) 	:	dataset dataset.name=? status.name=?
0.77%  (4363) 	:	dataset release.name=?
0.75%  (4257) 	:	file dataset.name=? run.run_number=?
0.68%  (3874) 	:	run run.run_number=?
0.53%  (2994) 	:	site site.name=?
0.45%  (2576) 	:	site file.name=?
0.45%  (2556) 	:	dataset file.name=?
0.43%  (2438) 	:	lumi file.name=?
0.40%  (2282) 	:	dataset dataset.name=? site.name=?
0.35%  (1999) 	:	file block.name=?
0.35%  (1970) 	:	dataset dataset.name=? run.run_number=?
0.29%  (1640) 	:	group dataset.name=?
0.29%  (1631) 	:	lumi run.run_number=?
0.25%  (1402) 	:	child dataset.name=?
0.22%  (1268) 	:	run file.name=?
0.21%  (1204) 	:	block block.name=?
0.19%  (1088) 	:	release release.name=?
0.16%  (936) 	:	site block.name=?
0.12%  (703) 	:	file dataset.name=? lumi.number=? run.run_number=?
0.10%  (594) 	:	parent file.name=?
0.08%  (455) 	:	dataset block.name=?
0.06%  (365) 	:	dataset dataset.name=? datatype.name=? release.name=?
0.06%  (360) 	:	file dataset.name=? site.name=?

Interesting non-valid queries:
T!_CERN
*herwig*/AODSIM
--> keyword search
lumi file = (file dataset=/RelValProdTTbar/JobRobot-MC_42_V12_JobRobot-v1/GEN-SIM-RECO)
file,lumi dataset=/RelValProdTTbar/JobRobot-MC_42_V12_JobRobot-v1/GEN-SIM-RECO
--> Users may like more complex combined queries
file dataset=/MuEG/Run2011B-PromptReco-v1/AOD, file.size >1
file dataset=/MinimumBias/Run2010A-valskim-v6/RAW-RECO* | grep run between  [138923, 144086] 
--> Users mixing up the post and pre filters
\end{verbatim}
% 
% Possible keyword search test cases:
% run duration run_number between [195100,195200] 
%
% py dataset=/DYJetsToLL_M-50_TuneZ2Star_8TeV-madgraph-tarball/Summer12_DR53X-PU_S10_START53_V7A-v1/AODSIM
%dataset dataset=*Run2011* datatype=data dataset=*RAW*

\appendix
\section{Data providers statistics}
\begin{verbatim}
(Some of the largest ones)

DBS:
    DB  size: 80GB + 200GB indexes, not many changes to existing  (old) records
    Largest tables: 
      Dataset (164K rows) -> Block (2M) -> Files (31M) -> FileRunLumi (902M) <- Runs (65K)

Phedex: ~7GB, more often changes to existing (even old) records
    change rate: 2,359,934 file transfers last month (from site A to site B; 
                 change rate on the DB to be found out)
\end{verbatim}

### 2012-10-11 doc/thesis/latex/literature_review.tex
updated problem definition, solutions and literature review. review to be finished

\textbf{\color{red}TODO: Overview; evaluate performance}

\subsection{Searching 'Deep Web' and Heterogeneous web services}

In 1990s there were was much research on search based service integration systems, for instance \textit{Information Manifold} following Local-as-View approach and \textit{TSIMMIS} developed at Stanford following Global-as-View approach. The \textit{Information Manifold} is quite resembling to DAS {\color{red},  so it could be useful checking it's papers. TODO: finish}

\subsubsection*{Deep web search at Google}
[multiple papers from Google] discusses various ways for implementing data integration in terms of large-scale search engine (Google) Virtual integration vs. surfacing. They also present ways for integrating systems without human intervention through use of statistical 'mediator'.

@@

\textit{Runtime query reformulation} - 'leaves data at the sources and routes queries to appropriate services'\cite[p. 1]{paygo_integration}

\textit{Deep-web surfacing} - tries to add content from the deep-web into search index. There are algorithms which allow to iteratively choose input parameters to forms to surface most of the 'hidden' data without loosing much (i.e. if choosing parameters in not smart way the web form could yield as many results as a cross product of all input combinations).

PayGo approach - there is NO single mediated schema over which users pose queries. Queries are routed to to the relevant sources. Statistical methods used to model uncertainty at all levels: queries, mappings and underlying data.

\subsubsection*{Then there is no access to index data terms}
{\color{red} TODO: Describe Keymantic\cite{Keymantic10}. This is a very suboptimal solution as indexing terms improves results (mention evaluation from SODA paper). It works only then keywords map entity names.

some hybrid approaches: 
index if exists
regexp
some string similarity measure based on historical data (e.g. even edit-distance would work for many items, like site)}

\subsection{Keyword search over DBs}
The problem of Keyword search over Relational Databases (or also semi-structured sources like XML) has received a significant attention by the research community over the last decade.

@@

\subsubsection*{Meta-data approach}

With a goal to bridge the increasing gap between high-level (conceptual, business) and low level (physical) representations of data, researchers from \textit{ETHZ} have been investigating Generation of SQL for Business users at \textit{Credit Suisse}.  For converting natural language queries 
% (that in addition to keyword search could convey some semantic structure)
 into SQL statements, in addition to what used by earlier approaches they used meta-data describing the schema (at multiple representation levels) and the domain (ontologies) and some natural language processing.

Even on a large data-warehouse of ~220GB data with a complex schema of 400+ tables they reported that if good meta-data is available, generating even quite complex SQL  (n-way joins with aggregations, etc) is quite feasible for computer. That would making it 'much easier for business users to interactively explore highly-complex data warehouses' \cite[p.932]{ethz2012}.

>>>>

\subsection{Overview}

The problem of keyword search over \textit{relational and other structured databases} is widely covered,  including even keyword search directly over distributed relational databases\cite[ch.16]{principles_data_integration} and some approaches for automatic schema matching has been discussed also. Such search is however very ambiguous and has quite large search space over the possible join paths, therefore most approaches rely on returning the top-k results.

Regarding the Boolean retrieval that we are the most interested, keyword query could be translated into list of best structured queries (e.g. SODA system proposing SQL over complex schemas, while Query Forms would propose mappings from keywords to parameters in SQL templates, Keymantic tries to achieve this without having access to Data itself). 
% TODO: Keymantic would propose SQL queries knowing only schema but no index

%  TODO: Further, {\color{red} in theory,} Google's Deep-web crawl techniques initially designed for accessing content hidden behind html forms could be used to effectively index the contents even of  data providers that do not provide a specific interface to do so.

{\color{red}Current research trends seem to fall on methods of automatic integration (e.g. statistical mediator), and pay-as-you-go integration. (TODO: see future chapter of \cite{principles_data_integration})}

%\textbf{\color{red}TODO: Overview; evaluate performance}

\subsection{Searching 'Deep Web' and web services}

In 1990s there were was much research on search based service integration systems, for instance \textit{Information Manifold} quite resembling DAS {\color{red},  so it could be useful checking it's papers.}.
% following Local-as-View approach
%and \textit{TSIMMIS} developed at Stanford following Global-as-View approach. The \textit{Information Manifold} is

\subsubsection*{No access to index data terms}

\cite{Keymantic10, semantics_without_access} explores the case then there is no possibility to index the data terms, e.g. then a DB is behind a wrapper (e.g. accessible only through a \textit{Web form} in “Hidden Web” or \textit{a web-service}) then crawling is generally not possible.

In Keymantic\cite{Keymantic10} a 'keyword query is processed as follows: First, all keywords that correspond to metadata items (e.g., field names) are extracted. The remaining
keywords are considered as possible input fields. Second, the likelihood of a remaining keyword to a metadata item is computed in order to rank different options to execute the keyword query on the
“Hidden Web” database'\cite[p.942]{ethz2012}.

\subsubsection*{Deep web search at Google}
[multiple papers from Google] discusses various ways for implementing data integration in terms of large-scale search engine (Google): Virtual integration vs. Surfacing. They also present ways for integrating systems without human intervention through use of statistical 'mediator'.

@@

\textit{Runtime query reformulation} - 'leaves data at the sources and routes queries to appropriate services'\cite[p. 1]{webscale_paygo}

\textit{Deep-web surfacing} - tries to add content from the deep-web into search index. There are algorithms which allow to iteratively choose input parameters to forms to surface a considerable part of the 'hidden' data without large overhead%
	\footnote{i.e. if choosing parameters in not smart way, a web form with just a couple of free inputs (or even dropdowns that's easier case), could yield as many results as a cross product of all input combinations.}.

\textit{PayGo approach}\cite{webscale_paygo}  - there is NO single mediated schema over which users pose queries. Queries are routed to to the relevant sources with help statistical methods that are used to model uncertainty at all levels: queries, mappings and underlying data.

%{\color{red} TODO: Describe Keymantic\cite{Keymantic10}. This is a very suboptimal solution as indexing terms improves results (mention evaluation from SODA paper). It works only then keywords map entity names. some hybrid approaches: 
%index if exists, regexp, some string similarity measure based on historical data (e.g. even edit-distance would work for many items, like site)}

\subsection{Keyword search over relational Database}
The problem of Keyword search over Relational Databases (or also semi-structured sources like XML) has received a significant attention by the research community over the last decade.

@@

\subsubsection*{SODA: Meta-data approach}

With a goal to bridge the increasing gap between high-level (conceptual, business) and low level (physical) representations of data, researchers from \textit{ETHZ} have been investigating Generation of SQL for Business users over a very complex data warehouse at \textit{Credit Suisse}.  For converting natural language queries 
% (that in addition to keyword search could convey some semantic structure)
 into SQL statements, in addition to what used by earlier approaches they used meta-data describing the schema at both physical, conceptual and logical levels extended with DBpedia (for synonyms, etc) and domain ontologies (to capture business concepts like 'wealthy customer').
%and some natural language processing

Even on a large data-warehouse of ~220GB data with a complex schema of 400+ tables they reported that if good meta-data is available, generating even fairly complex SQL  (e.g. n-way joins with aggregations) is quite feasible for a computer. That would make it 'much easier for business users to interactively explore highly-complex data warehouses' \cite[p.932]{ethz2012}. The users also reported system's potential a) for analysing the schema and learning patterns about it and b) as tool to help documenting legacy systems.

\subsection{Keyword search: integration on demand}
{\color{red}TODO?: \cite[ch.16]{principles_data_integration}}

### 2012-10-11 doc/thesis/latex/proposed_solutions.tex
updated problem definition, solutions and literature review. review to be finished

\begin{itemize}
\item Interactive guidance on how to compose queries to ease the learning (could also include some examples of popular queries):
\begin{enumerate}
\item What entity are you searching for?
\item How could you identify it (i.e. what do you know)?
\item After seeing first results: What do you want to do with the results? (list items; select only specific columns; do aggregation; use in command line/another program)
\end{enumerate}

\item (?) JavaScript-based query interpreter to ease query writing: could suggest search attribute and entity attribute names. also taking into account some possible ambiguous namings (auto completion)

\item Keyword query --> ranked list of matching structured queries:
	\begin{itemize}
	\item map keywords to entity name and selection parameters (quite resembling 'Query forms' and 'Keyword cleaning' approaches)
	\item
		{\color{red} Inverted index could be built with help of some good full-text search engine (e.g. Xapian\footnote{http://xapian.org/}) and used to generate mapping to structured queries. If no match found (e.g. new entry not yet in our cache) some pattern matching could be used to try to corresponding guess search attribute}
	\item ranking could be based on: how closely keywords are matching some or all required parameters, popularity of certain query and users feedback

	\item {\color{red}	--> evaluate different approaches (HMM, etc); could some natural language processing help?}
	\item special attention to \textbf{wilcard} queries

	\end{itemize}
\end{itemize}

\subsection{Performance}
\subsubsection{Continuous view maintenance at Data providers with large DBs}
Use either \textit{materialized refresh fast views with query rewriting} (Oracle; completely transparent for proprietary apps)\cite{Oracle11}
 or some other continuous view maintenance tool (e.g. DBToaster\footnote{http://www.dbtoaster.org that is being developed at EPFL}) to improve performance of heavy queries containing joins and/or aggregations.

\subsubsection{More intelligent caching}
Some of the data entities instances changing very rarely, for example in DBS system old datasets would never change, while new ones are constantly added (still some of their attributes may change, in this example validity).

Cached copy may be shown before hand while up to date results could be retrieved on user's request.

An automatic change rate prediction could be useful to efficiently balance between caching and retrieving results.

Also consider:
                			\begin{itemize}
                			\item given an entity received from provider, determining if it is useful to cache for long-term
%	                			\item \textbf{\color{red}At what level are we caching now? query or individual query results?}
                			\item could cache even \textit{expired} data, but warning user
                			\item could also have different validity dates for certain fields. if no volatile fields are not explicitly requested, even a very old cache could be used. 
                			\item Can we automatically figure which fields are static and which are changing?
                			\item pre-fetching common (sub-)queries: determining manually and/or automatically

                			\item \textbf{Deciding if to cache or not...}

@@

>>>>

\begin{itemize}
\item Interactive guidance on how to compose queries to ease the learning (could also include some examples of popular queries):
\begin{enumerate}
\item What entity are you searching for?
\item How could you identify it (i.e. what do you know)?
\item What do you want to do with these results (after seeing the results)? Possible options: list items; select only specific columns; do aggregation; use in command line/another program
\end{enumerate}

\item Given a keyword query, suggest the best matching structured queries (as a ranked list)%
 	\footnote{As DAS uses Boolean IR model we could only rank specific structured queries, but not the results itself.}:
	\begin{itemize}
	\item map keywords to entity names and selection attributes (quite resembling 'Query forms' and 'Keyword cleaning' approaches)
	\item build an inverted index of attribute values with some good full-text search toolkit\footnote{e.g. Xapian (http://xapian.org/) implemented in C++ which that is known for its flexibility, or Sphinx (http://sphinxsearch.com/)}. If no match found (e.g. new entry not yet in our cache) pattern matching could be used to make a best guess.
	\item ranking could be based on: how closely keywords are matching (some or all) required API's input parameters, popularity of certain queries and users feedback

	\item {\color{red} evaluate different user feedback approaches (e.g. HMM)
	\item multi-wildcard queries (e.g. dataset=*Zmm*special*RECO*) are common at CMS that will need special attention\footnote{Matching to structured query will be harder and guessing even worse. Also for performance reasons some web-services will introduce limitations on the format of wilcard queries, while DAS wants still to be flexible at least for datasets}.}

\item making use of direct access to database {\color{red}(if available; in theory)}
\begin{itemize}
\item one could do incremental indexing of a target database  \footnote{E.g. an inverted index of values in Oracle/MySQL DB tables can be built to keep the list of possible terms. For instance, Sphinx full text search engine can access DBs directly and also supports incremental indexing but it needs a bit of manual configuration {http://sphinxsearch.com/docs/current.html\#delta-updates} }, then try to map database table columns into API parameters (manually or even automatically by applying some of Data Integration approaches)
\item then this index + mapping could be used to improve the keyword search (better mapping from keywords to API inputs) or even structured search (e.g. query cleaning/term auto-completion)
\item problem: the owners of services may not want to provide direct access to DB because of no trust, security or performance issues
\item may need to exclude very large tables
\end{itemize}
	\end{itemize}
\item (?) JavaScript-based query interpreter to ease query writing: could suggest search attribute and entity attribute names. also taking into account some possible ambiguous namings (auto completion)

@@

\subsection{Performance}

%%%%%  -------------------------------------------

\subsubsection*{More intelligent caching}
Some of the data entities instances changing very rarely, for example in DBS system old datasets would never change, while new ones are constantly added (still some of their attributes may change, in this example validity). 
%
Cached copy may be shown by default while up to date results could be retrieved on user's request. An automatic change rate prediction could be useful to efficiently balance between caching and retrieving results.

On the other hand some result items may not make sense to be cached, as there may be too many of valid input parameter combinations (e.g. now there are 900M of \textit{(run, file)} combinations).

Also consider:
                			\begin{itemize}
                			%\item given an entity received from provider, determining if it is useful to cache for long-term
%	                			\item \textbf{\color{red}At what level are we caching now? query or individual query results?}
                			%\item could cache even \textit{expired} data, but warning user
                			\item could also have different validity dates for certain fields. if no volatile fields are not explicitly requested, even a very old cache could be used. 
                			\item Can we automatically figure which fields are static and which are changing?
                			\item pre-fetching common (sub-)queries: determining manually and/or automatically
                			\end{itemize}

%%%%%  -------------------------------------------
\subsubsection*{Continuous view maintenance at Data providers with large DBs}
In the cases when new records are coming not the existing ones are not changing much, continuous view maintenance that computes only differences from earlier results could be a good solution to improve the performance.

Use either \textit{materialized refresh fast views with query rewriting} (Oracle; completely transparent for proprietary apps)\cite{Oracle11}
 or some other continuous view maintenance tool (e.g. DBToaster\footnote{http://www.dbtoaster.org that is being developed at EPFL}) to improve performance of heavy queries containing joins and/or aggregations.

For example currently in DBS (80GB + 280GB indexes) to process a query 'find files where run in [r1, r2, r3] and dataset=X' one needs to join all the tables below.

\begin{verbatim}
Dataset (164K rows) -> Block (2M) -> Files (31M) -> FileRunLumi (902M) <- Runs (65K)
\end{verbatim}

However having a materialized view of all these tables joined together would allow answering such queries much quicker. Given low change rates (in comparison to data already present), maintaining the view should be also comparatively cheap with only expense of just couple of times of space.

%%%%%  -------------------------------------------
